### 2020-03-03 Papers/Torque_Paper/article_IOP.tex
final Torque Paper as submitted

\documentclass[a4paper]{jpconf}

%\journal{ }

% packages
%\usepackage{microtype}
%\usepackage{algorithm}
%\usepackage[noend]{algpseudocode}
%\usepackage[a4paper, total={6.2in, 9in}]{geometry}
%\renewcommand{\baselinestretch}{1.2} 
%\usepackage{color}
%\usepackage[x11names]{xcolor}
%\usepackage[utf8]{inputenc}
%\usepackage{amsmath}
%\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsfonts}
%\usepackage{tikz}
% \usepackage{array}
\usepackage{mathtools}
\usepackage{relsize}
%\usepackage[T1]{fontenc}
\usepackage{cases}
\usepackage{amsthm}
\usepackage{subcaption}
\usepackage{multirow}
%\usepackage{stix2}
%\usepackage[labelfont=bf,font=footnotesize]{caption}
\usepackage{tabularx,ragged2e,booktabs}
\usepackage{todonotes}
%\usepackage{lineno}
%\modulolinenumbers[5]
%\usepackage[toc]{appendix}
\usepackage{placeins}
\usepackage{hyperref}

%\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}
\newcommand{\lnorm}{\left|\left|}
\newcommand{\rnorm}{\right|\right|}
\newcommand{\Ld}{L^2(\mathcal{D})}
\newcommand{\Lomtd}{L^2(\Omega;\mathcal{T};\mathcal{D})}
%\newcommand{\Lomd}{L^2(\Omega;L^2(\mathcal{D}))}
\newcommand{\Lomd}{L^2(\Omega,\mathcal{D})}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

% bibliography
\newcommand{\myreferences}{../references,../Mendeley_refs}
\bibliographystyle{iopart-num} %elsarticle-num}

\numberwithin{equation}{section}
\graphicspath{{./images/}}

%\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}

\pagestyle{myheadings}

%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%\author{Jacky Mucklow}

%\address{Production Editor, \jpcs, \iopp, Dirac House, Temple Back, Bristol BS1~6BE, UK}

%\ead{jacky.mucklow@iop.org}

\title{Global sensitivity analysis of model uncertainty in aeroelastic wind turbine models}

%$^{\mathsf{1}}$ Centrum Wiskunde \& Informatica, Science Park 123, Amsterdam, The Netherlands.

\author{P.~Kumar$^{1}$, B.~Sanderse$^{1}$, K.~Boorsma$ ^{2}$, M.~Caboni$^{2}$} \ead{b.sanderse@cwi.nl}
\address{$^{1}$ Centrum Wiskunde \& Informatica (CWI), Amsterdam, The Netherlands}
\address{$^{2}$ TNO, Westerduinweg 3, Petten, The Netherlands}

%\author{Benjamin~Sanderse}
%\textsl{•}

\begin{abstract}
A framework is presented for performing global sensitivity analysis of model parameters associated with the Blade Element Momentum (BEM) models.  Sobol indices based on adaptive sparse polynomial expansions are used as a measure of global sensitivities. The sensitivity analysis workflow is developed using the uncertainty quantification toolbox \texttt{UQLab} that is integrated with TNO's \texttt{Aero-Module} aeroelastic code. Uncertainties in chord, twist, and lift- and drag-coefficients have been parametrized through the use of NURBS curves. Sensitivity studies are performed on the NM80 wind turbine model from the DanAero project, for a case with 19 uncertainties in both model and geometry. The combination of parametrization and sparse adaptive polynomial chaos yields a new efficient framework for global sensitivity analysis of aeroelastic wind turbine models, paving the way to effective model calibration. 
\end{abstract}

\vspace{-1cm}
%\begin{keyword}
%BEM, global sensitivity analysis, model uncertainty, Sobol indices
%\end{keyword}
%\end{frontmatter}

%\linenumbers

\input{content}

\section*{Acknowledgements}
The research was sponsored by the Topsector Energy Subsidy from the Dutch Ministry of Economic Affairs and Climate.

%\newpage
\section*{References}
\bibliography{\myreferences}

%\newpage

%\appendix
%\input{literature_overview}

\end{document}